### 2015-12-20 paper/images/flow.tex
add paper

% Define block styles
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
text width=4em, text centered, rounded corners, minimum height=3em]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, minimum height=2em, node distance = 2.5cm]

\begin{figure}[h]
  \centering
  \begin{tikzpicture}[node distance = 1.5cm, auto]
    % Place nodes
    \node [block] (reomodel) {REO Models};
    \node [cloud, left of=reomodel] (human) {Human};
    \node [block, below left = of reomodel] (tsamodel) {TSA};
    \node [block, below right = of reomodel] (mealymodel) {Mealy Machines};
    %\node [cloud, left = of tsamodel] (verification) {Verification};
    %\node [cloud, right = of mealymodel] (blackbox) {Blackbox};
    % Draw edges
    \path [line] (reomodel) -- (tsamodel);
    \path [line] (reomodel) -- (mealymodel);
    %\path [line] (tsamodel) -- (evaluate);
    \path [line] (mealymodel) -- (tsamodel);
    \path [line] (mealymodel) |-  (reomodel);
    %\path [line] (update) |- (tsamodel);
    %\path [line] (decide) -- node {no}(stop);
    %\path [line,dashed] (human) -- (reomodel);
  \end{tikzpicture}
  \caption{Our Idea}
  \label{fig:idea}
\end{figure}